---
title: "likert chart"
output:
  html_document: default
  pdf_document: default
date: '2022-03-28'
---

## import library
```{r setup likert}
library(likert)
```

## Import data

```{r}
rawdata <- read_csv("~/data.csv")
```

##delete first and second row
```{r data cleansing}
data <- data[-c(1:2),]
```

## Create Age columns

```{r create age columns}
data$age <- 2022 - as.numeric(data$birthyear)
```

## Create Age groups

```{r create age groups}
data["age_group"] = cut(data$age, c(18, 44, 49, 64, Inf), c("18-44", "45-49", "50-64", ">64"), include.lowest=TRUE)
```

## Extract columns from data

```{r}
whenready <- data[c("age_group","whenready_1","whenready_2","whenready_3","whenready_4","whenready_5","whenready_6","whenready_7","whenready_8","whenready_9","whenready_10")]
```

## create levels

```{r create levels}
levels = c("I am already ready or doing it","In 1-4 months","In 4-6 months","In more than 6 months","When a vaccine will be ready or the virus will have disappeared","I don’t foresee a time when I will be ready to do it again","Not applicable for me")
```

##  Change Likert scores to factor and specify levels

```{r change likert scores}
whenready$whenready_1 = factor(whenready$whenready_1,levels,ordered = TRUE)
whenready$whenready_2 = factor(whenready$whenready_2,levels,ordered = TRUE)
whenready$whenready_3 = factor(whenready$whenready_3,levels,ordered = TRUE)
whenready$whenready_4 = factor(whenready$whenready_4,levels,ordered = TRUE)
whenready$whenready_5 = factor(whenready$whenready_5,levels,ordered = TRUE)
whenready$whenready_6 = factor(whenready$whenready_6,levels,ordered = TRUE)
whenready$whenready_7 = factor(whenready$whenready_7,levels,ordered = TRUE)
whenready$whenready_8 = factor(whenready$whenready_8,levels,ordered = TRUE)
whenready$whenready_9 = factor(whenready$whenready_9,levels,ordered = TRUE)
whenready$whenready_10 = factor(whenready$whenready_10,levels,ordered = TRUE)
```

## Create likert plot

```{r create likert plot}
whenready_plot <- whenready[c("whenready_1","whenready_2","whenready_4","whenready_4","whenready_5","whenready_6","whenready_7","whenready_8","whenready_9","whenready_10")]
colnames(whenready_plot) <- c("Travel domestically by plane for business or essential purposes","Travel internationally by plane for business or essential purposes","Travel domestically by plane for leisure","Travel internationally by plane for leisure","Go shopping in-person at stores or malls","Return to the workplace","Go eat in restaurants","Go to bars, clubs, or crowded spaces","Attend large public events like a festival or outdoor gathering","Take public transit")
results <- likert(as.data.frame(whenready_plot))
title1 <- "Perceptions of post-covid behaviours"
summary(results)
plot<- plot(results)
final_plot <- plot + ggtitle(title1)
final_plot
```
## Create likert plot by age groups and separate into two groups of question themes - travel and local activity

```{r create likert plots by groups}
colnames(whenready) <- c("age_group","Travel domestically by plane for business or essential purposes","Travel internationally by plane for business or essential purposes","Travel domestically by plane for leisure","Travel internationally by plane for leisure","Go shopping in-person at stores or malls","Return to the workplace","Go eat in restaurants","Go to bars, clubs, or crowded spaces","Attend large public events like a festival or outdoor gathering","Take public transit")
cluster_plot_travel <- plot(likert(as.data.frame(whenready[,2:5]),grouping = whenready$age_group)) + ggtitle("Perceptions of post-covid behaviours - travel")
cluster_plot_travel
cluster_plot_activity <- plot(likert(as.data.frame(whenready[,6:11]),grouping = whenready$age_group)) + ggtitle("Perceptions of post-covid behaviours - local activity")
cluster_plot_activity
```
## Interpretation from the graphs
As we can see from the first graph above (Perceptions of post-covid behaviours - travel), the majority of respondents were still being hesitated in any kind of travels due to the pandemic. With younger are the age, more optimistic they are when facing the travelling situation. Also, for the second graph above (Perceptions of post-covid behaviours - local activity), it is obviously shown that most of the respondents are reopen to go eat in restaurants and go shopping in-person at stores or malls, especially for latter one, even older group of respondents (65+) also start to doing that activity.However, for the local activities such as attend large public events, go to bars or clubs and take public transport are still under consideration. Generally speaking, younger age groups (18-44 and 45-49) are quite optimistic in terms of returning to the workplace.


